---
title: "Untitled"
format: gfm
---

## Part 1: Vectorizing Code
The following functions can be written to be more efficient without using parallel computing:

1. This function generates a n x k dataset with all its entries distributed Poisson with mean lambda.

```{r}
fun1 <- function(n = 100, k = 4, lambda = 4) {
  x <- NULL
  
  for (i in 1:n)
    x <- rbind(x, rpois(k, lambda))
  
  return(x)
}

fun1alt <- function(n = 100, k = 4, lambda = 4) {
  x <- sapply(1:n, \(x) rpois(k,lambda))
  return(x)
}

# Benchmarking
bench::mark(
  fun1(),
  fun1alt(), relative = TRUE, check = FALSE
)
```

2. Like before, speed up the following functions (it is OK to use StackOverflow)

```{r}
# Total row sums
fun1 <- function(mat) {
  n <- nrow(mat)
  ans <- double(n) 
  for (i in 1:n) {
    ans[i] <- sum(mat[i, ])
  }
  ans
}

fun1alt <- function(mat) {
  x <- rowSums(mat)
}

# Cumulative sum by row
fun2 <- function(mat) {
  n <- nrow(mat)
  k <- ncol(mat)
  ans <- mat
  for (i in 1:n) {
    for (j in 2:k) {
      ans[i,j] <- mat[i, j] + ans[i, j - 1]
    }
  }
  ans
}

fun2alt <- function(mat) {
  n <- nrow(mat)
  x <- sapply(1:n, \(x) cumsum(mat[x,])) |> t()
}

# Use the data with this code
set.seed(2315)
dat <- matrix(rnorm(200 * 100), nrow = 200)

# Test for the first
bench::mark(
  fun1(dat),
  fun1alt(dat), relative = TRUE
)

# Test for the second
bench::mark(
  fun2(dat),
  fun2alt(dat), relative = TRUE
)
```

3. Find the column max (hint: Check out the function max.col()).

```{r}
# Data Generating Process (10 x 10,000 matrix)
set.seed(1234)
x <- matrix(rnorm(1e4), nrow=10)

# Find each column's max value
fun2 <- function(x) {
  apply(x, 2, max)
}

fun2alt <- function(x) {
  sapply(1:ncol(x), \(k) max(x[,k]))
}

# Benchmarking
bench::mark(
  fun2(x),
  fun2alt(x), relative = TRUE
)
```

## Part 2: Rcpp code
As we saw in the Rcpp week, vectorization may not be the best solution. For this part, you must write a function using Rcpp that implements the propensity score matching algorithm. You can use Week 5’s lab as a starting point for the problem. Your C++ file should look something like the following:

```{r}
#include<Rcpp.h>

# using namespace Rcpp;
# 
# List psmatch(
#   NumericVector pscores,
#   LogicalVector is_treated
# )
# {
#   /*... setup the problem creating the output...*/
# 
#   /*
#   ... Implement your matching (start from Week 5's lab)... 
#   ... You have to consider that matches are done againts groups, i.e.,
#       Treated (is_treated == true) must be matched to control 
#       (is_treated == false)  
#   */
# 
#   // Returning
#   return List::create(
#     _["match_id"] = /*...*/
#     _["match_pscore"] = /*...*/,
#   );
# 
# }
```

