---
title: "Homework 2"
format: gfm
---

## Part 1: Vectorizing Code
The following functions can be written to be more efficient without using parallel computing:

1. This function generates a n x k dataset with all its entries distributed Poisson with mean lambda.

```{r}
fun1 <- function(n = 100, k = 4, lambda = 4) {
  x <- NULL
  
  for (i in 1:n)
    x <- rbind(x, rpois(k, lambda))
  
  return(x)
}

fun1alt <- function(n = 100, k = 4, lambda = 4) {
  # Draw randomly from a Poisson(lambda) n*k times
  rpois(n*k, lambda) |> matrix(nrow = 100, byrow = TRUE)
}

# Benchmarking
bench::mark(
  fun1(),
  fun1alt(), relative = TRUE, check = FALSE
)
```

2. Like before, speed up the following functions (it is OK to use StackOverflow)

```{r}
# Total row sums
fun1 <- function(mat) {
  n <- nrow(mat)
  ans <- double(n) 
  for (i in 1:n) {
    ans[i] <- sum(mat[i, ])
  }
  ans
}

fun1alt <- function(mat) {
  x <- rowSums(mat)
}

# Cumulative sum by row
fun2 <- function(mat) {
  n <- nrow(mat)
  k <- ncol(mat)
  ans <- mat
  for (i in 1:n) {
    for (j in 2:k) {
      ans[i,j] <- mat[i, j] + ans[i, j - 1]
    }
  }
  ans
}

fun2alt <- function(mat) {
  apply(dat,1,cumsum) |> t()
}

# Use the data with this code
set.seed(2315)
dat <- matrix(rnorm(200 * 100), nrow = 200)

# Test for the first
bench::mark(
  fun1(dat),
  fun1alt(dat), relative = TRUE
)

# Test for the second
bench::mark(
  fun2(dat),
  fun2alt(dat), relative = TRUE
)
```

3. Find the column max (hint: Check out the function max.col()).

```{r}
# Data Generating Process (10 x 10,000 matrix)
set.seed(1234)
x <- matrix(rnorm(1e4), nrow=10)

# Find each column's max value
fun2 <- function(x) {
  apply(x, 2, max)
}

fun2alt <- function(x) {
  colMax <- max.col(t(x))
  n <- ncol(x)
  out <-  rep(NA, n)
  for(i in 1:n){
    out[i] <- x[colMax[i],i]
  }
  out
}

# Benchmarking
bench::mark(
  fun2(x),
  fun2alt(x), relative = TRUE
)
```

## Part 2: Rcpp code
As we saw in the Rcpp week, vectorization may not be the best solution. For this part, you must write a function using Rcpp that implements the propensity score matching algorithm. You can use Week 5’s lab as a starting point for the problem. Your C++ file should look something like the following:

```{Rcpp}
#include <Rcpp.h>

using namespace Rcpp;

// [[Rcpp::export]]
 List psmatch(
   NumericVector pscores,
   LogicalVector is_treated
 )
 {
  /*... setup the problem creating the output...*/
  int n = static_cast<int>(is_treated.size());
  IntegerVector indices(n); 
  NumericVector values(n); 

  /*
  ... Implement your matching (start from Week 5's lab)...
  ... You have to consider that matches are done againts groups, i.e.,
      Treated (is_treated == true) must be matched to control
      (is_treated == false)
  */
  for (int i = 0; i < n; ++i) {
    // For each treated person, we want to find a matching control
    if (is_treated[i] == true) {
      
      int best_n = 0;
      double best_diff = std::numeric_limits < double >::max();
      
      for (int j = 0; j < n; ++j) {
      // controls being matched be ith treated subject to identify optimal match
      if (is_treated[j] == false) {
        
      // If it is lower, then update
      double diff = std::abs(pscores[i] - pscores[j]);
      if (diff < best_diff) {
        
        best_n = j;
        best_diff = diff;
        }
      }
      }

      values[i] = pscores[best_n]; // noting the matched prop score
      indices[i] = best_n + 1; // noting the matched control index, adjusted for R
    
}
    }

  // Returning
  return List::create(
    _["match_id"] = indices,
    _["match_pscore"] = values 
  );

 }
```

# Example
```{r}
set.seed(324)
x <- runif(25) # pscore
y <- rbinom(25,1,0.4) |> as.logical() # is_treated
ans <- psmatch(x,y)

cbind.data.frame(pscores=x,is_treated=y,ans$match_id, ans$match_pscore)
```

